---
title: "Join, as DB Tables"
author: "SunnyBingoMe"
date: "2016-06-29"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Prepare Data
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(data.table)
hero = "
name,       alignment, gender,   publisher
Magneto,    bad,       male,     MarvelDuplicate
Storm,      good,      female,   MarvelDuplicate
Batman,     good,      male,     DC
Joker,      bad,       male,     DC
Catwoman,   bad,       female,   DC
Hellboy,    good,      male,     Dark Horse Comics
"
hero = read_csv(hero, trim_ws = TRUE, skip = 1)
hero = data.table(hero)

publisher = "
publisher,   yr_founded
DC,              1934
MarvelDuplicate, 1939
MarvelDuplicate, 8888
Image,           1992
"
publisher = read_csv(publisher, trim_ws = TRUE, skip = 1)
publisher = data.table(publisher)
```

__For all the `join` commands, if a key is set for both dt, then `on = 'key_col_name'` can be elided.__

WARN: row orders are not guaranteed

-----------

# Left-Join & Inner-Join
left-join will keep all left-side table rows, including no-matched rows

__by default, nomatch cells := NA__
```{r, eval=FALSE}
hero[publisher, on = "publisher"]
```
```{r, echo=FALSE, cache=TRUE}
t = hero[publisher, on = "publisher"]
knitr::kable(t)
```

__nomatch = 0 means delete no-matched rows__ (becomes inner-join)
```{r, eval=FALSE}
hero[publisher, nomatch = 0, on = "publisher"]
```
```{r, echo=FALSE, cache=TRUE}
t = hero[publisher, nomatch = 0, on = "publisher"]
knitr::kable(t)
```


-----------

# Right-Join & Inner, all = TRUE
right-join will keep all right-side table rows, unfortunately, we have to __swap two data.table's__ to make a right-join (using left-join format)
```{r, eval=FALSE}
publisher[hero, on = "publisher"]
```
```{r, echo=FALSE, cache=TRUE}
t = publisher[hero, on = "publisher"]
knitr::kable(t)
```

__nomatch = 0 means delete no-matched rows__ (becomes inner-join again, with row & col orders may diff from the left-join's inner-join)


-----------

# Outer-/Full-Join & Inner-Join

Outer-(or full-) join will keep both tables content, with extra NA's

we need function `merge`
```{r, eval=FALSE}
merge(hero, publisher, by = "publisher", all = TRUE)
```
```{r, echo=FALSE, cache=TRUE}
t = merge(hero, publisher, by = "publisher", all = TRUE)
knitr::kable(t)
```

if not `all` content is kept, it becomes inner-join, again.
```{r, eval=FALSE}
merge(hero, publisher, by = "publisher")
```
```{r, echo=FALSE, cache=TRUE}
t = merge(hero, publisher, by = "publisher")
knitr::kable(t)
```

-----------

# Semi-Join (Only Keep Half Side Content)
```{r, cache=TRUE, results='hide'}
index = unique(hero[publisher, on = "publisher", which=TRUE, allow.cartesian=TRUE])
hero[index]
```
```{r, echo=FALSE, cache=TRUE}
t = hero[index]
knitr::kable(t)
```

you can add `nomatch = 0`

-----------

# Anti-Join

-----------

# Cross-Join

-----------
ref:

[data](http://stat545.com/bit001_dplyr-cheatsheet.html)

[semi-Join](http://stackoverflow.com/questions/18969420/perform-a-semi-join-with-data-table)

[SQL Join, chart explained](http://stackoverflow.com/questions/5706437/whats-the-difference-between-inner-join-left-join-right-join-and-full-join)

```{r}
devtools::session_info()
```

