---
title: "Get Started With data.table (DT)"
author: "SunnyBingoMe"
date: "2016-06-29"
output: 
    html_document: 
        toc: true
# output: github_document
# output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

See also [Notes of Practical Machine Learning - Coursera](http://tdd.nu/pml-notes) for ML and ggplot related commands.

The official "Getting Started" is [here](https://github.com/Rdatatable/data.table/wiki/Getting-started).

# DT Key Ideas
1. `data.table` is an extension of `data.frame`, it is capable of df
2. DT's syntax is similar to SQL: `DT[i, j, by]` = sql query `[where, select, group_by]`
i.e. `[WHERE_condition_of_rows, SELECT_of_columns, GROUP_BY_a_categorical_variable]`

# DT Basic Commands

## prepare data for commands
```{r init, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
```

```{r load data, message=FALSE, warning=FALSE, cache=TRUE}
system.time((dt = fread("D:/hNow/Dropbox/Github/r_data_table_start/gb_code.csv")))
summary(dt)
head(dt)
```

## subset
```{r subset rows, cache=TRUE}
# subset rows
subset_rows = dt[V4 == "England" & V3 == "Beswick" & V10 < 53.93]
subset_rows
```

```{r subset cols, cache=TRUE}
# subset cols
subset_cols = dt[ , .(V2,V4)]
head(subset_cols)
nrow(subset_cols)
```

wrong example using ` : `
```{r subset cols wrong, eval=FALSE}
subset_cols = dt[ , .(V2:V4)] # wrong !!!
```

## order/sort
```{r order, cache=TRUE}
dt_order = dt[order(V4, -V8)] # radix sort, so it is fast
```

## column actions
add, update, delete, etc.
tip: ` := `

### add (or unconditional update of all rows in a column):
```{r col add, cache=TRUE}
dt[ , V18 := V10 + V11]
head(dt)

dt[ , V19 := 'this-is-NEW.col']
head(dt)

system.time(
dt[ , V20 := paste(V18, V19, sep = "_")] # expensive, system.time = 16s. see also 'OBS' below
)
head(dt)
```

### update (with WHERE condition):
```{r col update}
dt[V8 == "Aberdeen City", V8 := "updated city name"]
head(dt)
```

### delete:
```{r col del}
# delete one col
dt[ , V18 := NULL]
head(dt)

# delete multi cols (must use their names)
dt[ , c('V19', 'V20') := NULL]
head(dt)
```

wrong format of delete:
```{r col del wrong, eval=FALSE}
dt[ , .(V11, V12) := NULL] # wrong !!!
dt[ , c(V11, V12) := NULL] # wrong !!!
```

### combine all tasks (just need to put `[]`s together):
this is called "chaining" [official doc](https://rawgit.com/wiki/Rdatatable/data.table/vignettes/datatable-keys-fast-subset.html)
```{r}
dt[V8 == "Aberdeen City", V8 := "updated_city"][ , V_New := V10 + V11][ , c("V6","V7") := NULL]
head(dt)
```

# DT Aggregate
Aggregation computes functions while grouping.
```{r}
summary(dt$V_New)
unique(dt$V4)

# calculate a "func"
dt[V_New >= 50, mean(V10), by = V4]

# named calculation in "func", instead of auto-named as 'V1', 'V2' ...
# OBS:  .()  is necessary, otherwise confusion with arg of dt
dt[V_New >= 50, .(averaged_result = mean(V10)), by = V4]

# multi-calculation
dt[V_New >= 50, .(sum(V10), mean(V10), .N), by = V4]

# multi-calculation with names in "func" and func in "by"
dt[V_New >= 50, .(sumTotal = sum(V10), avr = mean(V10), countNr = .N), by = substr(V4, 1,1)]
```

# Performance DT vs dplyr
```{r}
# data.table
system.time((
dt[V_New >= 50, .(sumTotal = sum(V10), avr = mean(V10), countNr = .N), by = substr(V4, 1,1)]
))

# base::aggregate
system.time({
mainDataDf = subset(dt, V_New >= 50)
summary_t = as.data.frame(as.list(aggregate(
    cbind(V10)
    ~ substr(V4, 1,1)
    ,data = mainDataDf
    ,FUN=function(mainDataDf) c(sumTotal = sum(mainDataDf), avr = mean(mainDataDf), countNr = length(mainDataDf))
)))
})
summary_t

# dplyr::summarise
system.time({
mainDataDf = subset(dt, V_New >= 50)
mainDataDf$t = substr(mainDataDf$V4, 1,1)
summary_t = summarise(group_by(mainDataDf, t),
                      sumTotal = sum(V10),
                      avr = mean(V10),
                      countNr = n())
mainDataDf$t = NULL
})
summary_t

# pip-ed dplyr::summarise
system.time({
summary_t = dt %>%
    filter(V_New >= 50) %>%
    mutate(t = substr(V4, 1,1)) %>%
    group_by(t) %>%
    summarise(sumTotal = sum(V10),
              avr = mean(V10),
              countNr = n())
})
summary_t
```

Note: the results are the same, but rwos in different order.
To make exactly the same, we can use `base::order()` and `dplyr::arrange()`.

## io performance
* `dt = fread("data.csv")` 10x faster than `read.csv("data.csv")`.
* `fwrite()` is faster than `write.csv()`. 

# DT Other Useful Features
* automatic indexing
* rolling joins
* overlapping range joins 
* and a lot more ...


# OBS
## cat/paste in `[]`
`paste` works fine.
wrong usage of `cat`:
```{r cat disaster, eval=FALSE}
system.time(dt[ , V20 := cat(V18, V19)])  # wrong !!!
# this will print all cat-results in console, and NOT adding/updating any column
```

# Ref
[ref](https://www.evernote.com/shard/s144/sh/4b76e1d6-20e0-476c-8c40-aaf32a62a75e/b9f375de89da3229)
